A comparative study of batch-construction strategies for recurrent neural networks in MXNet

### 2017-04-18 paper/paper.tex
paper

  \section{Bucketing in MXNet} \label{sec:bucketing}
  Bucketing, borrowed from TensorFlow's sequence training example, is a simple solution. Instead of unrolling the network to the maximum possible sequence, we unroll multiple instances of different lengths (e.g., length 5, 10, 20, 30). During training, we use the most appropriate unrolled model for each mini-batch of data with different sequence length. For RNNs, although those models have different architecture, the parameters are shared in time. So although models in different buckets are selected to train in different mini-batches, essentially the same set of parameters are being optimized. MXNet reuses the internal memory buffers among all executors.

  \section{Proposed Approach} \label{sec:approach}
  At each iteration:
  \begin{enumerate}
  	\item shuffle training data by sequence length
  	\item partition resulting sequence order into $N$ partitions
  	\item sort each bin $n$:
  	\begin{itemize}
  		\item in increasing order if $n$ is odd
  		\item in decreasing order if $n$ is even
  	\end{itemize}
  \end{enumerate}

  \section{Experimental Setup} \label{sec:setup}
  The 4th CHiME Speech Separation and Recognition Challenge
  \cite{Vincent_CSL2016:CHiME4} revisits the CHiME-3 dataset \cite{Barker2015:CHiME3},
  which consists of utterances spoken by speakers in challenging noisy environments recorded using a 6-channel microphone array on a tablet.
  We extracted 16-dimensional MFCC vectors from beam-formed versions of the six sub-corpora in the training data corresponding to each channel
  and used them as features for the neural network models \cite{menne16:chime4System}.
  A CART tying was used to cluster the (roughly 120k possible) allophones into 1500 CART labels.
  We trained a three state HMM without skip transitions in the standard fashion and applied it in a forced alignment mode to the training data
  in order to obtain a frame-wise labeling of the data.

  \section{Experiments} \label{sec:experiments}
   We conducted experiments with different batch construction strategies. The results are reported in \ref{tab:chime:batch}, where the first three rows show results for trivial sequence ordering 
   methods. The last two rows then provide a direct comparison of the bucketing approach as 
   it is implemented in MXNet and the Laplace approach as proposed in this paper.
    %\begin{table}[tbp]
    %\centering
    %\caption{Comparison of runtime and memory requirement for different software packages. The %numbers were averaged over 100 training epochs. 
    %Note that sharp memory usage estimates for Torch and TensorFlow can not be obtained due to %their 5internal memory management.}
    %\label{tab:chime:compare}
    %\begin{tabular}{lrr}
    %  \hline
    %  Toolkit                 & Runtime [sec] & Memory [GB] \\
    %  \hline
    %  RETURNN   &  140     & 0.6     \\
    %  \hline
    %  Theano (in RETURNN)     &  370     & 0.8     \\
    %  TensorFlow              &  580     & 2.1$^*$     \\
    %  \hline
    %\end{tabular}
    %\end{table}

@@

	\begin{table}[tbp]
		\centering
		\caption{An evaluation of different sequence batch construction methods on the CHiME-4 database. Training time per epoch, memory consumption are presented in the first two columns, while the last columns shows the word error rate of the corresponding system
		when decoded with the RASR speech recognition toolkit.}
		\label{tab:chime:batch}
		\begin{tabular}{lrrr}
			\hline
			Approach                & Runtime [sec] & Memory [GB] & WER [\%] \\
			\hline
			Static					&				&			  &			 \\
			Random 					&				&			  & 		 \\
			Sorted					& 				&			  &			 \\
			\hline
			Bucketing				&				&			  &			 \\
			Laplace					&				&			  &			 \\
			\hline
		\end{tabular}
	\end{table}

>>>>

  \section{Bucketing in MXNet} \label{sec:bucketing}
  Borrowed from TensorFlow's sequence training example, MXNet implements bucketing by 
  clustering sequences into bins depending on their length. The size of each bin, i.e.~ the span of 
  sequence lengths associated with this bin, has to be specified by the user and optimal values 
  depend on the ASR task. The sampling process can be done in logarithmic time, since for each sequence length in the training set a binary search over the bins can be performed.

  In each iteration of the mini-batch training a bucket is then selected randomly. Within the selected bucket a random span of sequences is chosen to be used as data batch. Note that this random shuffling only ensures a large inter-batch variance w.r.t.~the sequence length, while the variance 
  within each batch is small.

  Bucketing is especially suited if the RNN model itself is not able to handle arbitrary 
  long sequences but instead requires to store an unrolled version of the network for every possible 
  length. In those cases bucketing allows the framework to assign each batch to the shortest possible 
  unrolled network, while still optimizing the same shared weights.

  \section{Proposed Approach} \label{sec:approach}

  At each iteration:
  \begin{enumerate}
  	\item shuffle training data by sequence length
  	\item partition resulting sequence order into $N$ partitions
  	\item sort each bin $n$:
  	\begin{itemize}
  		\item in increasing order if $n$ is odd
  		\item in decreasing order if $n$ is even
  	\end{itemize}
  \end{enumerate}

  \section{Experimental Setup} \label{sec:setup}
  The 4th CHiME Speech Separation and Recognition Challenge
  \cite{Vincent_CSL2016:CHiME4} consists of noisy utterances spoken by speakers in challenging 
  environments. Recorded was done using a 6-channel microphone array on a tablet. The dataset revisits 
  the CHiME-3 corpus that was published one year before the challenge took place \cite{Barker2015:CHiME3}.

  We extracted 16-dimensional MFCC vectors from beam-formed versions of the six sub-corpora in the training data corresponding to each channel
  and used them as features for the neural network models \cite{menne16:chime4System}.
  A CART tying was used to cluster the (roughly 120k possible) allophones into 1500 CART labels.
  We trained a three state HMM without skip transitions in the standard fashion and applied it in a forced alignment mode to the training data
  in order to obtain a frame-wise labeling of the data.

  \section{Experiments} \label{sec:experiments}
   We conducted experiments with different batch construction strategies. The results are reported in \ref{tab:chime:batch}, where the first three rows show results for trivial sequence ordering 
   methods. The last two rows then provide a direct comparison of the bucketing approach as 
   it is implemented in MXNet and the sinusoidal approach as proposed in this paper.
    %\begin{table}[tbp]
    %\centering
    %\caption{Comparison of runtime and memory requirement for different software packages. The %numbers were averaged over 100 training epochs. 
    %Note that sharp memory usage estimates for Torch and TensorFlow can not be obtained due to %their 5internal memory management.}
    %\label{tab:chime:compare}
    %\begin{tabular}{lrr}
    %  \hline
    %  Toolkit                 & Runtime [sec] & Memory [GB] \\
    %  \hline
    %  RETURNN   &  140     & 0.6     \\
    %  \hline
    %  Theano (in RETURNN)     &  370     & 0.8     \\
    %  TensorFlow              &  580     & 2.1$^*$     \\
    %  \hline
    %\end{tabular}
    %\end{table}

@@

	\begin{table}[tbp]
		\centering
		\caption{An evaluation of different sequence batch construction methods on the CHiME-4 database. Training time per epoch, memory consumption are presented in the first two columns, while the last columns shows the word error rate of the corresponding system
		when decoded with the RASR speech recognition toolkit.}
		\label{tab:chime:batch}
		\begin{tabular}{lrrr}
			\hline
			Approach                & Runtime [sec] & Memory [GB] & WER [\%] \\
			\hline
			Static					&				&			  &			 \\
			Random 					&				&			  & 		 \\
			Sorted					& 				&			  &			 \\
			\hline
			Bucketing				&				&			  &			 \\
			Sinusoidal  			&				&			  &			 \\
			\hline
		\end{tabular}
	\end{table}
